---
title: "R Notebook"
output: html_notebook
---

Let's load in the data first and have a good look at it before working out exactly what needs to be cleaned.

```{r}
library(tidyverse)
library(readr)
library(here)

dirty_decathlon_data <- read_rds(here::here("raw_data/decathlon.rds"))
                                 
dirty_decathlon_data
```

```{r}
names(dirty_decathlon_data)
```


```{r} 
#Need to check this as there seems to be an error, could be important when renaming the columns later. Are there spaces at the end of most variable names?
stopifnot(
  names(dirty_decathlon_data) == c("100m", "Long.jump", "Shot.put", "High.jump", "400m", "110m.hurdle", "Discus", "Pole.vault", "Javeline", "1500m", "Rank", "Points", "Competition")
)

```


```{r}
view(dirty_decathlon_data)
```

```{r}
glimpse(dirty_decathlon_data)
```

Cool, looks like it's not too big a data set but this may be lulling us into a false sense of security. Let's take a look at the number of NAs across the board.

```{r}
ddd_NAs <-
dirty_decathlon_data %>%
summarise(across(.fns = ~sum(is.na(.x))))

ddd_NAs
```

Ok so as suspicious as it seems, NAs don't seem to be the issue here. The data seems to have read in correctly so this is clearly not the issue with this data set. At first glance, the following issues need to be addressed.

Prescribed in task:

1. This data has row names - how do you get rid of them?
2. This data isn’t tidy
3. You will want better column header names.

Others noticed:

1. The row names are of representative of competitors. There is some duplication, with some competitors having competed in two competitions (Decaster and Olympic G), details contained in it's own column.
2. The competitors are named for one competition with in ALL CAPS and for the other in lower case. 
3. The data contains a mixture of different measurements for different events (time and distance). Think that this just needs to be cleaned up in the column names. One extra is that times for the 1500m would usually be expressed a minute and second format, as would anything above 60 seconds, but maybe it's best to leave this or make sure that it only effects output?

Analysis questions that need to be answered:

1. Who had the longest long jump seen in the data?
2. What was the average 100m time in each competition?
3. Who had the highest total points across both competitions?
4. What was the shot-put scores for the top three competitors in each competition?
5. What was the average points for competitors who ran the 400m in less than 50 seconds     vs. those than ran 400m in more than 50 seconds?

So in terms of these questions, 

1. We'll have to make sure that we match the competitors who participated in each competition and that those that only participated in one competition do not influence things like averages.
2. Probably won't get onto this because it's being a little bit pernickety, but if we are to go a little bit further with Q3 and consider that it is supposed to be representative of **performance** across the two events, it's noticeable that the latter competition has more points applied to the same ranks. Maybe weighting could be used to counteract this effect, although it may well be that the systems awards greater points for more esteemed events.


**Plan of Action**

1. First and foremost, we need work out how to get rid of these row headings and to create a "Competitor column with cleaned names (i.e no ALL CAPS).
2. It's a tough call to say whether or not the data should stay long format or should be made wide format. I'm guessing it should stay as it is, although I'm not sure whether or not it would be better to have the data arranged by competitor or left in the current order of ranking and competition. 
3. Once this decision has been made, I'd say we need to rename the column headers in line with our house style (**STRIKE THAT, SHOULD DO THIS SECOND**).  
4. Could verify that each variable is within reasonable limits with a function but it hardly seems worth it considering the size of the data set. 

Let's make a start on these and then move from there:

1.

```{r}
# Searched help files for "Tools for working with row names"

ddd_rows_heads_rem <-
rownames_to_column(dirty_decathlon_data, var = "competitor")

ddd_rows_heads_rem

```
```{r}
ddd_comptr_no_caps <-
  ddd_rows_heads_rem %>%
  mutate(competitor = str_to_title(competitor))

ddd_comptr_no_caps

```

2. 

```{r}
#Decided to rename columns individually so that I could get everything lower case, with underscore spacing and with metres or seconds added.

ddd_clean_col_names <-
ddd_comptr_no_caps %>%
  rename(
    "100m_(sec)" = "100m",
    "long_jump_(m)" = "Long.jump",
    "shot_putt_(m)" = "Shot.put",
    "high_jump_(m)" = "High.jump",
    "400m_(sec)" = "400m",
    "100m_hurdle_(sec)" = "110m.hurdle",
    "discus_(m)" = "Discus",
    "pole_vault_(m)" = "Pole.vault",
    "javeline_(m)" = "Javeline",
    "1500m_(sec)" = "1500m",
    rank = Rank,
    points = Points,
    competition = Competition
  )

ddd_clean_col_names

```

```{r}
ddd_clean_arrange <-
ddd_clean_col_names %>%
  arrange(desc(competition), rank)

ddd_clean_arrange
```
```{r}
ddd_clean_arrange <-
ddd_clean_arrange %>%
  select(competitor, competition, rank, points, everything())  

ddd_clean_arrange
```

Think that should be good enough for us to start on the questions.

```{r}

```



