---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)

candy_data <- read_csv(here::here("clean_data/clean_candy_script.csv"))
```

Q1. What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)

```{r}
candy_data %>%
  summarise(ratings_count = sum(!is.na(candy_data$rating)))
  
```


Q2. What was the average age of people who are going out trick or treating and the average age of people 3. not going trick or treating?

```{r}
#Checking the NAs
candy_data %>%
  summarise(age_NAs = sum(is.na(candy_data$age)))
```

```{r}
#Two different ways of getting the answer without NAs
candy_data %>%
  filter(!is.na(age), (!is.na(going_out))) %>%
  group_by(going_out) %>% 
  summarise(round(mean(age)))
```

```{r}
candy_data %>%
  drop_na(age, going_out) %>%
  group_by(going_out) %>%
  summarise(round(mean(age)))
```

```{r}
#Calculating the mean with medians in place of NAs. 
candy_data %>%
  drop_na(going_out) %>%
  mutate(age = coalesce(age, median(age, na.rm = TRUE))) %>%
  group_by(going_out) %>%
  summarise(round(mean(age)))
```

Q3. For each of joy, despair and meh, which candy bar revived the most of these ratings?

```{r} 
#Currently doing all these questions with certain candies not put together (i.e, x100_grand_bar = 100_grand_bar. Can add put these together easily in the code if need be later and this analysis should still work, but with potentially different results)

candy_rankings <-
candy_data %>%
  drop_na(rating) %>%
  group_by(candy_type) %>% 
  summarise(
    joy = sum(rating == "JOY"),
    despair = sum(rating == "DESPAIR"),
    meh = sum(rating == "MEH")
    )

candy_rankings
```

```{r}
candy_rankings %>%
  slice_max(joy)
  
candy_rankings %>%
  slice_max(despair) 

candy_rankings %>%
  slice_max(meh)

```

```{r}
# Any full-size candy bar seems quite generic, probably the reason for the number of positive responses. Maybe better to check the best rated specific candy

candy_rankings %>%
  arrange(desc(joy)) %>%
  slice(2)

```


Q4. How many people rated Starburst as despair?

```{r}
candy_rankings %>% 
  filter(candy_type == "starburst") %>%
  select(candy_type, despair)
```

For the next three questions, count despair as -1, joy as +1 and meh as 0.

Q6. What was the most popular candy bar by this rating system for each gender in the dataset?



Q7. What was the most popular candy bar in each year?

Q8. What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?
