---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)

candy_data <- read_csv(here::here("clean_data/clean_candy_data.csv"))
```

```{r}
candy_data %>%
distinct(gender)
```



Q1. What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)

```{r}
candy_data %>%
  summarise(ratings_count = sum(!is.na(candy_data$rating)))
  
```


Q2. What was the average age of people who are going out trick or treating and the average age of people 3. not going trick or treating?

```{r}
#Checking the NAs
candy_data %>%
  summarise(age_NAs = sum(is.na(candy_data$age)))
```

```{r}
#Two different ways of getting the answer without NAs
candy_data %>%
  filter(!is.na(age), (!is.na(going_out))) %>%
  group_by(going_out) %>% 
  summarise(round(mean(age)))
```

```{r}
candy_data %>%
  drop_na(age, going_out) %>%
  group_by(going_out) %>%
  summarise(round(mean(age)))
```

```{r}
#Calculating the mean with medians in place of NAs. 
candy_data %>%
  drop_na(going_out) %>%
  mutate(age = coalesce(age, median(age, na.rm = TRUE))) %>%
  group_by(going_out) %>%
  summarise(round(mean(age)))
```

Q3. For each of joy, despair and meh, which candy bar revived the most of these ratings?

```{r} 
#Currently doing all these questions with certain candies not put together (i.e, x100_grand_bar = 100_grand_bar. Can add put these together easily in the code if need be later and this analysis should still work, but with potentially different results)

candy_rankings <-
candy_data %>%
  drop_na(rating) %>%
  group_by(candy_type) %>% 
  summarise(
    joy = sum(rating == "JOY"),
    despair = sum(rating == "DESPAIR"),
    meh = sum(rating == "MEH")
    )

candy_rankings
```

```{r}
candy_rankings %>%
  slice_max(joy)
  
candy_rankings %>%
  slice_max(despair) 

candy_rankings %>%
  slice_max(meh)

```

```{r}
# Any full-size candy bar seems quite generic, probably the reason for the number of positive responses. Maybe better to check the best rated specific candy

candy_rankings %>%
  arrange(desc(joy)) %>%
  slice(2)

```


Q4. How many people rated Starburst as despair?

```{r}
candy_rankings %>% 
  filter(candy_type == "starburst") %>%
  select(candy_type, despair)
```

For the next three questions, count despair as -1, joy as +1 and meh as 0.

Q6. What was the most popular candy bar by this rating system for each gender in the dataset?

```{r}
num_rated_candy <-
candy_data %>% 
  drop_na(rating) %>% 
  mutate(score = case_when(
    rating == "JOY" ~ 1,
    rating == "DESPAIR" ~ (-1),
    rating == "MEH" ~ 0))

```


```{r}
gender_group <-
num_rated_candy %>%
  select(gender, candy_type, score) %>% 
  group_by(candy_type, gender) %>% 
  summarise(total_score = sum(score))

gender_group
```

```{r}
#Still getting "any_full_size_candy_bar" here so decided to take the top three until we have a specific type of candy for each
gender_group %>%
  group_by(gender) %>% 
  slice_max(total_score, n = 3)
```

Q7. What was the most popular candy bar in each year?

```{r}
year_group <-
num_rated_candy %>%
  select(year, candy_type, score) %>% 
  group_by(candy_type, year) %>% 
  summarise(total_score = sum(score))

year_group
```

```{r}
#Same again here
year_group %>%
  group_by(year) %>% 
  slice_max(total_score, n = 3)
```


Q8. What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?

```{r}
country_group <-
num_rated_candy %>%
  select(country, candy_type, score) %>% 
  group_by(candy_type, country) %>% 
  summarise(total_score = sum(score))

country_group
```

```{r}
country_group %>%
  group_by(country) %>% 
  slice_max(total_score, n = 3)
```

