---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)

decathlon_data <- read_csv(here::here("clean_data/clean_decathlon_data"))

```

```{r}
decathlon_data
```


Q1. Who had the longest long jump seen in the data?

```{r}
decathlon_data %>%
  arrange(
    desc(`long_jump_(m)`)
    ) %>% 
  slice_max(
    `long_jump_(m)`
    ) %>%
  select(
    competitor, `long_jump_(m)`
    )
```

Q2. What was the average 100m time in each competition?

```{r}
decathlon_data %>% 
  group_by(
    competition
    ) %>%
  summarise(
    avg_100m_time = mean(`100m_(sec)`)
    ) %>%
  arrange(avg_100m_time)

```


Q3. Who had the highest total points across both competitions?

```{r}
decathlon_data %>%
  group_by(
    competitor
    ) %>%
  summarise(
    total_points = sum(points)
    ) %>%
  slice_max(total_points)
```


Q4. What was the shot-put scores for the top three competitors in each competition?

```{r}
decathlon_data %>% 
  filter(
    rank <= 3
    ) %>%
  select(competitor, competition, rank, `shot_put_(m)`)
  
```

Q5. What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?

```{r}
decathlon_data %>%
  group_by("400m_under_50_secs" = `400m_(sec)` < 50) %>%
  summarise(avg_points = mean(points))

```

**End of Questions**